---
title: "Explore the Data"
description: "Data set description."
format:
  html:
    df-print: kable
    code-fold: true
    code-summary: "Show code"
    code-overflow: wrap
    toc-title: Page Contents
    toc: true
    toc-depth: 2
    toc-location: right
    number-sections: false
    html-math-method: katex
    css: styles.css
    theme: flatly
    smooth-scroll: true
editor_options: 
  chunk_output_type: console
---

# Overview

This page will show plots and figures describing the contents of the f(l)usion data set.

## Data description:

```{r warning=FALSE, message=FALSE}
get_fusion <- function(url) {
  library(readr)
  df <- read_csv(url)
  return(df)
}

hub_url <- "https://github.com/JMHumphreys/flusion/raw/main/flusion/flusion_v1.csv"

myFlusion <- as.data.frame(get_fusion(hub_url))


head(myFlusion)
```

#### Data attributes:

-   *date*: Date of the reported epidemiological week (epiweek) based on a Saturday start.\
-   *year*: Year of estimate\
-   *epiweek*: The epidemiological week\
-   *abbreviation*: U.S. State or Territory abbreviation.\
-   *location*: Numeric location code based on FIPS.\
-   *location_name*: Name of U.S. State or Territory.\
-   Columns beginning with *q\_* provide the 0.025, 0.25, 0.5, 0.75, and 0.975 quantiles for estimated flu hospitalizations.

## Plot National Totals

```{r fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
library(tidyverse)
options(dplyr.summarise.inform = FALSE, show_col_types = FALSE)

natl_sums <- myFlusion %>%
  group_by(date) %>%
  summarise(Q0.25 = sum(q_0.25),
            Q0.50 = sum(q_0.50),
            Q0.75 = sum(q_0.75))


ggplot(natl_sums, aes(date, Q0.50)) +  
  geom_ribbon(aes(ymin=Q0.25, ymax=Q0.75),fill="steelblue", alpha = 0.8) +
  geom_line(linewidth = 0.5) +
  scale_x_date(date_breaks = "6 month", date_labels = "%b-%Y") +
  ylab("Estimated Hospitalizations") +
  xlab(" ") +
  theme_minimal() +
  theme(plot.margin = unit(c(2,0.1,2,0.1), "cm"),
        panel.grid.minor = element_line(color = "gray90", linewidth = 0.25, linetype = 1),
        panel.grid.major = element_line(color = "gray60", linewidth = 0.5, linetype = 1),
        panel.background = element_blank(),
        plot.background = element_blank(),
        legend.position="none",
        legend.text = element_text(size=12, face="bold"),
        legend.title = element_text(size=16, face="bold"),
        axis.title.x =  element_text(size=16, face="bold"),
        axis.title.y = element_text(size=16, face="bold"),
        axis.text.x =  element_text(size=14, face="bold", angle=60, hjust=1),
        axis.text.y = element_text(size=12, face="bold"),
        plot.title = element_text(size=22, face="bold"))
```

## Plot States

```{r fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
set.seed(123)
random_states <- sample(myFlusion$abbreviation, size=4)

states_plot <- myFlusion %>%
  filter(abbreviation %in% random_states)

ggplot(states_plot, aes(date, q_0.50)) +  
  geom_ribbon(aes(ymin=q_0.025, ymax=q_0.975),fill="steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin=q_0.25, ymax=q_0.75),fill="steelblue", alpha = 0.5) +
  geom_line(linewidth = 0.5) +
  scale_x_date(date_breaks = "6 month", date_labels = "%b-%Y") +
  facet_grid(rows = vars(location_name), scales = "free_y") +
  ylab("Estimated Hospitalizations") +
  xlab(" ") +
  theme_minimal() +
  theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
        panel.grid.minor = element_line(color = "gray90", linewidth = 0.25, linetype = 1),
        panel.grid.major = element_line(color = "gray60", linewidth = 0.5, linetype = 1),
        panel.background = element_blank(),
        plot.background = element_blank(),
        strip.text = element_text(size=14, face="bold"),
        strip.background = element_blank(),
        legend.position="none",
        legend.title = element_text(size=16, face="bold"),
        axis.title.x =  element_text(size=16, face="bold"),
        axis.title.y = element_text(size=16, face="bold"),
        axis.text.x =  element_text(size=12, face="bold", angle=60, hjust=1),
        axis.text.y = element_text(size=12, face="bold"),
        plot.title = element_text(size=22, face="bold"))

```

## Short Term

```{r, warning=FALSE, message=FALSE}
set.seed(111)
random_states <- sample(myFlusion$abbreviation, size=4)

random_yr <- sample(myFlusion$year, size=1)



states_plot_yr <- myFlusion %>%
  filter(abbreviation %in% random_states,
         year == random_yr | year == (random_yr + 1))

ggplot(states_plot_yr, aes(date, q_0.50)) +  
  geom_ribbon(aes(ymin=q_0.025, ymax=q_0.975),fill="steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin=q_0.25, ymax=q_0.75),fill="steelblue", alpha = 0.5) +
  geom_line(linewidth = 0.5) +
  scale_x_date(date_breaks = "6 month", date_labels = "%b-%Y") +
  facet_grid(rows = vars(location_name), scales = "free_y") +
  ylab("Estimated Hospitalizations") +
  xlab(" ") +
  theme_minimal() +
  theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
        panel.grid.minor = element_line(color = "gray90", linewidth = 0.25, linetype = 1),
        panel.grid.major = element_line(color = "gray60", linewidth = 0.5, linetype = 1),
        panel.background = element_blank(),
        plot.background = element_blank(),
        strip.text = element_text(size=14, face="bold"),
        strip.background = element_blank(),
        legend.position="none",
        legend.title = element_text(size=16, face="bold"),
        axis.title.x =  element_text(size=16, face="bold"),
        axis.title.y = element_text(size=16, face="bold"),
        axis.text.x =  element_text(size=16, face="bold"),
        axis.text.y = element_text(size=12, face="bold"),
        plot.title = element_text(size=22, face="bold"))

```
